### 2016-03-08 ch05-initvals/initial-values.tex
Algorithm formatting improved

Consequently, $V(\theta)$ is a non-quadratic function of $\theta$ which, in general, results in a non-convex optimization problem. 
The Levenberg-Marquardt~\citep{Marquardt1963} algorithm (see Algorithm~\ref{LMalgo}) is used to solve this optimization problem deterministically.
Such an approach requires good initial estimates of $\theta$ to avoid inherent local optima, which is the focus of this chapter.

\begin{algorithm}
\caption{Levenberg-Marquardt~\citep{Marquardt1963}, \citep[Sec. 9.L.4]{Pintelon2012}}\label{LMalgo}
\begin{algorithmic}[1]
  \Require Cost function \eqref{eq:MLEcf} is rewritten as $V(\theta) = \InnerProductSelf{\epsilon(\theta)}$. 
\Function{LevenbergMarquardt}{$\epsilon(\theta),  \theta^{\mathrm{init}}$}
   \State $\atIter[0]{\theta} \gets \theta^{\mathrm{init}}$
   \Comment{$\atIter[i]{X}$ denotes $X$ at iteration $i$.}
   \For{$i$ \textbf{in} $1 \to \infty$}
      \State $U \, \Sigma \, W^{\TT} \gets \code{svd}\left( \frac{\partial \epsilon(\atIter[i-1]{\theta})}{\partial \theta}\right)$ 
        \Comment{Singular value decomposition of Jacobian}
      % \State $
      %                  \text{Denote }\Sigma \equiv 
      %                  \left[ \begin{matrix} 
      %                  \sigma_1         & \deemph{0} & \deemph{\cdots} &           & \deemph{0}      \\
      %                  \deemph{0}      & \sigma_2    &                   &           &                   \\
      %                  \deemph{\vdots} &              & \ddots            &           & \deemph{\vdots} \\
      %                                    &              &                   & \sigma_n &                   \\
      %                   \deemph{0}      &              & \deemph{\cdots} &           & \deemph{0}      
      %                   \end{matrix}\right]
      %                $
      % \State \textbf{if} $i=1$ \textbf{ then} $\lambda \gets \max(\sigma) / 100$
      \If{$\lambda$ is undefined}
        \State $\lambda \gets \max(\Set{\Sigma_{11}, \ldots, \Sigma_{nn}})/100$
      \EndIf
      \State $\Sigma_{kk} \gets \frac{\Sigma_{kk}}{(\Sigma_{kk}^2 + \lambda^2)} \quad \forall k \in \Set{1, \ldots, n}$
      \State $\delta\theta \gets - W  \Sigma^{\TT}  U^{\TT} \epsilon(\atIter[i-1]{\theta})$
      \State $\atIter[i]{\theta} \gets \frac{(\atIter[i-1]{\theta} + \delta\theta) }{ \norm[2]{\atIter[i-1]{\theta} + \delta\theta} }$
      \If{$V(\atIter[i]{\theta}) > V(\atIter[i-1]{\theta})$} \Comment{Current step is a deterioration}
         \State $\atIter[i]{\theta} \gets \atIter[i-1]{\theta}$ \Comment{Restart loop from previous estimate}
          \State $\lambda \gets 10 \lambda$ 
      \Else
          \State $\lambda \gets 0.4 \lambda$
      \EndIf
      \If {$\norm[2]{\delta\theta} < n_{\theta}  \cdot 10^{-6} \norm[2]{\atIter[i]{\theta}}$}
          \State \textbf{return} $\hat{\theta} \gets \atIter[i]{\theta}$
      \EndIf
   \EndFor
\EndFunction
\end{algorithmic}
\end{algorithm}

>>>>

Consequently, $V(\theta)$ is a non-quadratic function of $\theta$ which, in general, results in a non-convex optimization problem. 
The Levenberg-Marquardt algorithm~\citep{Marquardt1963}, is used to solve this optimization problem deterministically and is shown in~\algoref{alg:initvals:levenberg-marquardt}.
Such an approach requires good initial estimates of $\theta$ to avoid inherent local optima, which is the focus of this chapter.

\begin{algorithm}
\caption{Levenberg-Marquardt~\citep{Marquardt1963},~\citep[Sec. 9.L.4]{Pintelon2012}}
\label{alg:initvals:levenberg-marquardt}
\begin{algorithmic}[1]
  \Require Cost function \eqref{eq:MLEcf} is rewritten as $V(\theta) = \InnerProductSelf{\epsilon(\theta)}$. 
\Function{LevenbergMarquardt}{$\epsilon(\theta),  \theta^{\mathrm{init}}$}
   \State $\atIter[0]{\theta} \gets \theta^{\mathrm{init}}$
   \Comment{$\atIter[i]{X}$ denotes $X$ at iteration $i$.}
   \For{$i$ \textbf{in} $1 \to \infty$}
      \State $U \, \Sigma \, W^{\TT} \gets \code{svd}\left( \frac{\partial \epsilon(\atIter[i-1]{\theta})}{\partial \theta}\right)$ 
        \Comment{Singular value decomposition of Jacobian}
      % \State $
      %                  \text{Denote }\Sigma \equiv 
      %                  \left[ \begin{matrix} 
      %                  \sigma_1         & \deemph{0} & \deemph{\cdots} &           & \deemph{0}      \\
      %                  \deemph{0}      & \sigma_2    &                   &           &                   \\
      %                  \deemph{\vdots} &              & \ddots            &           & \deemph{\vdots} \\
      %                                    &              &                   & \sigma_n &                   \\
      %                   \deemph{0}      &              & \deemph{\cdots} &           & \deemph{0}      
      %                   \end{matrix}\right]
      %                $
      % \State \textbf{if} $i=1$ \textbf{ then} $\lambda \gets \max(\sigma) / 100$
      \If{$\lambda$ is undefined}
        \State $\lambda \gets \max(\Set{\Sigma_{11}, \ldots, \Sigma_{nn}})/100$
      \EndIf
      \State $\Sigma_{kk} \gets \frac{\Sigma_{kk}}{(\Sigma_{kk}^2 + \lambda^2)} \quad \forall k \in \Set{1, \ldots, n}$
      \State $\delta\theta \gets - W  \Sigma^{\TT}  U^{\TT} \epsilon(\atIter[i-1]{\theta})$
      \State $\atIter[i]{\theta} \gets \frac{(\atIter[i-1]{\theta} + \delta\theta) }{ \norm[2]{\atIter[i-1]{\theta} + \delta\theta} }$
      \If{$V(\atIter[i]{\theta}) > V(\atIter[i-1]{\theta})$} \Comment{Current step is a deterioration}
         \State $\atIter[i]{\theta} \gets \atIter[i-1]{\theta}$ \Comment{Restart loop from previous estimate}
          \State $\lambda \gets 10 \lambda$ 
      \Else
          \State $\lambda \gets 0.4 \lambda$
      \EndIf
      \If {$\norm[2]{\delta\theta} < n_{\theta}  \cdot 10^{-6} \norm[2]{\atIter[i]{\theta}}$}
          \State \textbf{return} $\hat{\theta} \gets \atIter[i]{\theta}$
      \EndIf
   \EndFor
\EndFunction
\end{algorithmic}
\end{algorithm}
